---
title: "Markdown In-Class Exercise"
author: "Jackson Dial"
date: "8/31/2021"
output: word_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = F,
	message = FALSE,
	warning = FALSE
)
```

```{r message=FALSE, warning=FALSE}
library(knitr)
library(flextable)
library(tidyverse)
```

## Introduction

This exercise will use the **CO2** dataset, which is part of base R. The CO2 dataset is taken from an experiment on the cold tolerance of the grass species _Echinochloa crus-galli_.

The CO2 dataset has 84 rows and 5 variables, including:

1. Plant ID
2. Plant origin
3. Treatment group
4. Ambient carbon dioxide concentration (mL/L)
5. CO2 uptake rate ( _umol/m_^2 sec)  


## Data Details

* There are 12 different plant IDs in this dataset

* There are 2 plant origins:
  + Quebec
  + Mississippi

* There are 2 treatment groups:
  + Chilled
  + Nonchilled

\pagebreak
  
## Results

The overall mean ambient carbon dioxide concentration was 435mL/L and the overall mean CO2 uptake rate was 27.21 _umol/m_^2 sec. Means by treatment and origin can be found in Table 1.

Making a table with kable:

```{r}
Names <- c("Quebec", "Mississippi", "chilled", "nonchilled")
Mean <- c(mean(CO2[CO2$Type == "Quebec",]$uptake),
               mean(CO2[CO2$Type == "Mississippi",]$uptake),
                    mean(CO2[CO2$Treatment == "chilled",]$uptake),
                         mean(CO2[CO2$Treatment == "nonchilled",]$uptake)

)

SD <- c(sd(CO2[CO2$Type == "Quebec",]$uptake),
               sd(CO2[CO2$Type == "Mississippi",]$uptake),
                    sd(CO2[CO2$Treatment == "chilled",]$uptake),
                         sd(CO2[CO2$Treatment == "nonchilled",]$uptake)

)




df <- cbind.data.frame(Names, Mean, SD)


kable(df, digits = 2, caption = "Table 1")
```

Making a table with flextable:

```{r}
ft = flextable(df)

# color(ft, ~ SD > 10, ~ SD, color = "red2")
# colformat_num(ft, j = c(2,3), digits = 2)


ft %>% color(~ SD > 10, ~ SD, color = "red2") %>% 
  colformat_double(j = c("Mean", "SD"), digits = 2) 
```

\pagebreak

The figure below shows the relationship between ambient CO2 concentration and CO2 uptake in the plants

```{r, echo = T}
library(ggplot2)

ggplot(data=CO2, aes(x=conc, y=uptake)) + geom_point(color=4, size=3) + labs(x="CO2 Concentration", y="CO2 Uptake", title="Figure 1. CO2 concentration and uptake") + theme_bw()

```





