---
title: "Project1_JacksonDial"
author: "Jackson Dial"
date: "9/3/2021"
output: word_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load Libraries

```{r}
#setwd("C:/Users/jacks/OneDrive/Desktop/DUKE_FALL2021/721/BIOS721")
#libraries
library(here)
library(readr)
library(tidyverse)

```


## Read Data

```{r}
#use here() function for better reproducibility
patient <- read_csv(here("Data", "Patient Level Data_proj1.csv"))
encounter <- read_csv(here("Data", "Encounter Level Data_proj1.csv"))

#join the patient level and encounter level data frames into one on MRN (ID)
dat <- encounter %>%
  left_join(patient, by = "MRN")

```

### Data Explanation

The data included in this analysis is comprised of two original data sets, one on patient data and one on encounter data. Patient data can be understood as information regarding a specific person, containing demographic information. This original dataset has 50 observations (patients) and 8 variables. The encounter dataset is a record of every time any of the 50 patients had an interaction with the Healthcare system being considered, and contains information regarding that visit such as reason for visit, date, and measurements regarding their current health status. This dataset contains 550 observations and 7 variables. A more precise description of each variable will be given below.

These two described data sets were joined together on the "MRN" column, which is an acronym meaning Medical Record Number. The dimension of this data frame is 550 observations with 14 variables. 






